---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# snvecR

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/snvecR)](https://CRAN.R-project.org/package=snvecR)
[![GPL-3](https://github.com/japhir/snvecR/blob/master/LICENSE.md)](https://img.shields.io/github/license/japhir/snvecR?logo=gnu&.svg)
[![release](https://img.shields.io/github/v/release/japhir/snvecR.svg)](https://img.shields.io/github/v/release/japhir/snvecR)
[![R-CMD-check](https://github.com/japhir/snvecR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/japhir/snvecR/actions/workflows/R-CMD-check.yaml)
[![Launch binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/japhir/snvecR/main)
<!-- badges: end -->

The goal of snvecR is to calculate precession and obliquity from an orbital
solution (defaults to ZB18a) and assumed or reconstructed values for tidal
dissipation (T<sub>d</sub>) and dynamical ellipticity (E<sub>d</sub>).

## Installation

You can install the development version of snvecR like so:

``` r
remotes::install_github("japhir/snvecR")
```

## Example

Here's the main function that does the work in action:

```{r example}
library(snvecR)
solution <- snvec()
```

see `?snvec` for further documentation.

Here we create a quick plot of the calculated climatic precession with the
eccentricity envelope:

```{r plot}
library(ggplot2)
solution |>
  ggplot(aes(x = age, y = cp)) +
  # the age scale goes from old to young
  scale_x_reverse() +
  # plot climatic precession
  geom_line(colour = "skyblue") +
  # add the (interpolated) eccentricity envelope
  geom_line(aes(y = eei))
```
