---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# snvecR

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/snvecR)](https://CRAN.R-project.org/package=snvecR)
<!-- badges: end -->

The goal of snvecR is to calculate precession and obliquity from an orbital solution (defaults to ZB18a) and assumed or reconstructed values for tidal dissipation (td) and dynamical ellipticity (ed).

## Installation

You can install the development version of snvecR like so:

``` r
remotes::install_github("japhir/snvecR")
```

## Example

Here's the main function that does the work in action:

```{r example}
library(snvecR)
solution <- snvec()
```

see `?snvec` for further documentation.

Here we create a quick plot of the calculated climatic precession with the eccentricity envelope:

```{r plot}
library(ggplot2)
solution |>
  ggplot(aes(x = age, y = cp)) +
  # the age scale goes from old to young
  scale_x_reverse() +
  # plot climatic precession
  geom_line() +
  # add the (interpolated) eccentricity envelope
  geom_line(aes(y = eei))
```
